---
title: Forecasting US Import
output:
  md_document:
    variant: gfm+footnotes
    preserve_yaml: TRUE
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../_posts") })
date: 2026-03-30
permalink: /posts/2026/forecast_us_import
excerpt_separator: <!--more-->
always_allow_html: false
toc: false
header:
tags:
  - tidyverse
  - data-science
  - visualization
  - data-manipulation
  - forecasting
  - trade
editor_options: 
  markdown: 
    wrap: 72
---

```{r packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(here)
library(forecast)
library(astsa)
library(lubridate)
library(zoo)
library(MLmetrics)

```

# Introduction

Here, I use US import data gathered from the USITC database. I
harmonized HTS codes across revisions using Pierce and Schott's (2012)
concordance. I originally cleaned data for a portion of my dissertation
that requires differentiation in import from Canada, Mexico, China, and
the Rest of the World (ROW). However, for this exercise, I'm going
aggregate all import values to annual value of consumable imports
(`con_val_mo`) and dutiable import for consumption (`dut_val_mo`).

```{r load data}
import_df <- readRDS(here("local", "us_monthly_import_df_1990_2016.hs_synth_df.rds")) %>%
  # get rid of negative import values 
  filter(dut_val_mo >= 0 & con_val_mo >= 0) %>%
  group_by(year, month) %>%
  summarise(con_val_mo = sum(con_val_mo, na.rm=T),
            dut_val_mo = sum(dut_val_mo, na.rm=T)) %>%
  ungroup()


```

Next, let's examine the trends across time. The data is at the monthly
level, which provides greater granularity to the data. Aside from the
two massive drop in early 1990s and 2000s, we saw a modest drop around
the time of the financial crisis.

```{r}
# Describe the data

import_df %>%
  # create year-month Date object using the lubridate package
  mutate(ym = paste0(year, "-", month, "-01"),
         ym = ymd(ym)) %>%
  ggplot(aes(x = ym, y = log(con_val_mo))) +
  geom_line() +
  scale_x_date(date_breaks = "4 year", date_labels = "%Y")
```

It looks like there were two massive drops in imports. This may be a
result from administrative data error? Let's take the 12-month rolling
average of the data.

```{r rolling average}
import_df %>%
  mutate(con_val_mo_rollavg = rollapply(con_val_mo, width = 12, FUN = mean, fill = NA, align = "right", partial = TRUE),
         dut_val_mo_rollavg = rollapply(dut_val_mo, width = 12, FUN = mean, fill = NA, align = "right", partial = TRUE)) %>%
  mutate(ym = paste0(year, "-", month, "-01"),
         ym = ymd(ym)) %>%
  ggplot(aes(x = ym, y = log(con_val_mo_rollavg))) +
  geom_line() +
  scale_x_date(date_breaks = "4 year", date_labels = "%Y")
```

Given the shock in 2008, I will conduct a forecasting analysis from
1993 - 2008, omitting 1992-09 and 2004-02 due to outlier import value.
Next, I split the data into a training and validation set, using
1990-2004 as training and 2005-2008 as validation.

```{r}
import <- import_df %>%
    mutate(ym = paste0(year, "-", month, "-01"),
         ym = ymd(ym))  %>%
  mutate(con_val_mo = ifelse(ym %in% c("1992-09-01", "2004-02-01"), NA, con_val_mo)) %>%
  # interpolate these years
  mutate(con_val_mo = na.approx(con_val_mo)) %>%
  select(ym, con_val_mo) 
  

validation <- import %>% 
  filter(year(ym) >= 2004) 

training_ts <- ts(
  import %>% filter(year(ym) < 2004) %>% pull(con_val_mo),
  frequency = 12, # Assuming monthly data based on your period2
  start = c(1990, 1) # Adjust to match your dataset's actual starting year and month
)

validation_ts <- ts(
  import %>% filter(year(ym) >= 2004 & year(ym) <= 2012) %>% pull(con_val_mo),
  frequency = 12,
  start = c(2004, 1)
)

```

```{r}
hw_model = hw(training_ts, period1 = 4, period2 = 12, h = length(validation_ts))

MAPE(hw_model$mean, validation_ts) *100


forecast_df <- tibble(
  # Assuming 'validation' is still in your environment from the previous step
  ym = validation$ym, 
  predicted_val = as.numeric(hw_model$mean)
)

import %>%
  ggplot(aes(x = ym, y = log(con_val_mo))) +
  geom_line(color = "black") + # Historical data
  
  # Add the predicted values over the top
  geom_line(data = forecast_df, 
            aes(x = ym, y = log(predicted_val)), 
            color = "red", 
            linetype = "dashed", 
            linewidth = 1) +
            
  scale_x_date(date_breaks = "4 year", date_labels = "%Y") +
  labs(y = "Log(Import Value)", x = "Date") +
  theme_minimal()
```

```{r}
arima_optimal = auto.arima(training_ts)

sarima_forecast = sarima.for(training_ts, n.ahead=length(validation_ts),
                             p=0,d=0,q=0,P=0,D=1,Q=0,S=12)

arima_forecast <- forecast(arima_optimal, h = length(validation_ts))

# 2. Extract the point forecasts into a tidy data frame 
# (Exactly like we did for the Holt-Winters model!)
arima_forecast_df <- tibble(
  ym = validation$ym, 
  predicted_val = as.numeric(arima_forecast$mean)
)

# 3. Plot it over your historical data
import %>%
  ggplot(aes(x = ym, y = log(con_val_mo))) +
  geom_line(color = "black") + 
  
  # Add the new ARIMA predictions
  geom_line(data = arima_forecast_df, 
            aes(x = ym, y = log(predicted_val)), 
            color = "blue", # Switched to blue to distinguish from HW!
            linetype = "dashed", 
            linewidth = 1) +
            
  scale_x_date(date_breaks = "4 year", date_labels = "%Y") +
  labs(y = "Log(Import Value)", x = "Date", title = "ARIMA Forecast") +
  theme_minimal()

```

# Reference

Pierce, Justin R. and Peter K. Schott. 2012, “Concording U.S. Harmonized
System Categories Over Time.” Journal of Official Statistics 28(1): pp.
53-68.
