---
title: POLI 144AB Coding Workshop 2
output:
  md_document:
    variant: gfm+footnotes
    preserve_yaml: TRUE
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../_posts") })
date: 2024-08-08
permalink: /posts/2024/poli144ab_coding2
excerpt_separator: <!--more-->
always_allow_html: true
toc: true
header:
tags:
  - tidyverse
  - data-science
  - visualization
  - data-manipulation
editor_options: 
  markdown: 
    wrap: 72
---

[R
Script](https://www.dropbox.com/scl/fi/51ru0k8p2gove2bl9iylf/workshop2.Rmd?rlkey=hjxvje7lgqvy5hpsn24bxap72&dl=0){:
.btn--research}

[V-Dem Dataset
(v9)](https://www.dropbox.com/scl/fi/0oioskebzm75q0bzdnwxk/V-Dem-Dataset.csv?rlkey=k9so4ygzoenan627aq00x4c3y&dl=0){:
.btn--research} [V-Dem Dataset
Codebook](https://www.dropbox.com/scl/fi/ffeimy5337k79x3t1frx2/V-Dem-Codebook-v9.pdf?rlkey=fmqxn3dsbjru8ag2oburlqklx&dl=0){:
.btn--research}

<!--more-->

**Dataset Citation:** Coppedge, Michael, John Gerring, Carl Henrik
Knutsen, Staffan I. Lindberg, Jan Teorell, David Altman, Michael
Bernhard, M. Steven Fish, Adam Glynn, Allen Hicken, Anna Lührmann, Kyle
L. Marquardt, Kelly McMann, Pamela Paxton, Daniel Pemstein, Brigitte
Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Agnes Cornell,
Lisa Gastaldi, Haakon Gjerløw, Valeriya Mechkova, Johannes von Römer,
Aksel Sundtröm, Eitan Tzelgov, Luca Uberti, Yi-tingWang, Tore Wig, and
Daniel Ziblatt. 2019. "V-Dem Codebook v9" Varieties of Democracy (V-Dem)
Project.

## Load data

```{r setup, include=FALSE}
# define paths
base.dir <- "~/Dropbox/Website/"
base.url <- "/"
fig.path <- "images/post/poli144ad_coding2/"

# this is where figures will be sent
paste0(base.dir, fig.path)

# this is where markdown will point for figures
paste0(base.url, fig.path)

knitr::opts_knit$set(base.dir = base.dir, base.url = base.url)

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      fig.path = fig.path,
                      fig.align = 'center',
                      fig.width = 6,
                      fig.height = 6,
                      dpi = 300)

```

```{r, message=FALSE, warning=FALSE}

library(here)
library(readr)
library(tidyverse)
library(haven)

vdem<- read_csv("~/Dropbox/Teaching/SS2_2024_POLI144AB/Data/V-Dem Dataset.csv")
```

## Select our variables

```{r, message=FALSE, warning=FALSE}
vdem_cleaned <- vdem %>% 
  select(country_name, year, v2x_polyarchy, v2mecenefm) 
```

## Histogram

```{r, message=FALSE, warning=FALSE}
vdem_cleaned %>%
  ggplot(aes(x = v2x_polyarchy)) +
  geom_histogram()

```

```{r, message=FALSE, warning=FALSE}
vdem_cleaned %>%
  ggplot(aes(x = year, y = v2x_polyarchy, color = country_name)) +
  geom_line() +
  theme(legend = "none") +
  guides(color="none")
# quite messy here!

# let's take a summary across all countries for each year
vdem_cleaned %>%
  group_by(year) %>%
  summarise(v2x_polyarchy_average = mean(v2x_polyarchy, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = v2x_polyarchy_average)) +
  geom_line()
# that's better!
```

## Scatterplot

```{r, message=FALSE, warning=FALSE}
vdem_cleaned %>%
  ggplot(aes(x = v2x_polyarchy, y = v2mecenefm)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r, message=FALSE, warning=FALSE}
lm(v2mecenefm ~ v2x_polyarchy,
   data = vdem_cleaned)
```

```{r, message=FALSE, warning=FALSE}
-1.575 + 4.479
```

```{r, message=FALSE, warning=FALSE}
vdem_global_mean <- vdem_cleaned %>%
  group_by(year) %>%
  summarise(v2x_polyarchy = mean(v2x_polyarchy, na.rm = TRUE)) %>%
  mutate(country_name = "Global")

vdem_cleaned %>%
  filter(country_name == "United States of America") %>%
  select(-v2mecenefm) %>%
  bind_rows(vdem_global_mean) %>%
  ggplot(aes(x = year, y = v2x_polyarchy, color = country_name)) +
  geom_line() +
  labs(x = "Year",
       y = "V-Dem Polyarchy",
       color = "Country",
       title = "Polyarchy Trends over Time (US vs Global)")


```

```{r, message=FALSE, warning=FALSE}

vdem_cleaned %>%
  filter(country_name %in% c("United States of America", "Mexico", "Canada")) %>%
  select(-v2mecenefm) %>%
  ggplot(aes(x = year, y = v2x_polyarchy, fill = country_name)) +
  geom_col(position = "dodge") +
  scale_fill_viridis_d()


```

```{r, message=FALSE, warning=FALSE}
library(plotly)

piechart_data = vdem_cleaned %>%
  filter(country_name %in% c("United States of America", "Mexico", "Canada") & year == 2000) 


plot_ly(piechart_data, labels = ~country_name, values = ~v2x_polyarchy, type = "pie", marker = list(colors = c("#F23030", "#267365", "#F28705")) )

```
