---
title: POLI 144AB Coding Workshop 3
output:
  md_document:
    variant: gfm+footnotes
    preserve_yaml: TRUE
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../_posts") })
date: 2024-08-13
permalink: /posts/2024/poli144ab_coding3
excerpt_separator: <!--more-->
always_allow_html: false
toc: false
header:
tags:
  - tidyverse
  - data-science
  - visualization
  - data-manipulation
editor_options: 
  markdown: 
    wrap: 72
---

[R
Script](https://www.dropbox.com/scl/fi/0sstwkux9dx3aw2gdjv2a/Workshop3.Rmd?rlkey=0wm7y03904vd12s74e130uojc&dl=0){:
.btn--research}

<!--more-->

```{r setup, include=FALSE}

# define paths
base.dir <- "~/Dropbox/Website/"
base.url <- "/"
fig.path <- "images/post/poli144ad_coding3/"

# this is where figures will be sent
paste0(base.dir, fig.path)

# this is where markdown will point for figures
paste0(base.url, fig.path)

knitr::opts_knit$set(base.dir = base.dir, base.url = base.url)

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      fig.path = fig.path,
                      fig.align = 'center',
                      fig.width = 6,
                      fig.height = 6,
                      dpi = 300)


```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Introduction

Linear regression is a fundamental statistical method used to model the
relationship between a dependent variable and one or more independent
variables. In this lesson, we will cover the basics of linear
regression, including how to implement it in R.

### What is Linear Regression?

Linear regression aims to find the best-fitting straight line through
the data points, predicting the dependent variable (Y) from the
independent variable(s) (X).

#### Simple Linear Regression

Simple linear regression involves one dependent variable and one
independent variable. The model is represented as:

$$Y = \beta_0 + \beta_1X + \epsilon$$

-   $Y$ is the dependent variable.
-   $X$ is the independent variable.
-   $\beta_0$ is the intercept.
-   $\beta_1$ is the slope of the line.
-   $\epsilon$ is the error term.

## Dataset

For this lesson, we'll use the `mtcars` dataset available in R.

```{r data, message=FALSE, warning=FALSE}
# Load the dataset
data(mtcars)
force(mtcars)

# Build the linear model
# lm(dependent variable ~ independent variable)
model_1 <- lm(mpg ~ wt, data = mtcars)
model_2 <- lm(mpg ~ hp, data = mtcars)

# Summary of the model
summary(model_1)
summary(model_2)

# Plotting the data points
mtcars %>%
  ggplot(aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = "lm", se = T)

# Build the multiple linear regression model
model_mult <- lm(mpg ~ wt + hp, data = mtcars)

# Summary of the model
summary(model_mult)

```

A one unit increase in HP (horse power) is correlated with 0.03177 units
lower in mile per gallon and is statistically significant at \< 0.001
level. A one unit increase in WT (weight) is correlated with 3.877 units
lower in mile per gallon and is statistically significant at 0.001
level.

## Generating Predictions

After building a linear regression model, we can use it to make
predictions. We'll cover how to generate predictions for both the simple
(bivariate) and multiple (multivariate) linear regression models.

### Predictions with Simple Linear Regression

Using the `model` we created earlier (predicting `mpg` based on `wt`),
let's predict the `mpg` for a car that weighs 3,000 lbs.

```{r prediction, message=FALSE, warning=FALSE}
new_data <- data.frame(wt = 3)

# Generate prediction

predicted_mpg <- predict(model_1, newdata = new_data)

# Output the prediction

predicted_mpg


ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue", se = FALSE) +
  geom_point(aes(x = new_data$wt, y = predicted_mpg), color = "red", size = 3) +
  geom_text(aes(x = new_data$wt, y = predicted_mpg, label = round(predicted_mpg, 2)), 
            vjust = -1, color = "red") +
  labs(title = "MPG vs Weight with Prediction",
       x = "Car Weight (1000 lbs)",
       y = "Miles Per Gallon") +
  theme_minimal()



```
